Digital Analysis of Paintings

### 2014-05-07 paper.tex
Add paper 2.

>>>>

Calculating the backbone is based on the edge detection of a segmented
brushstroke, this will generally form ``Y'' shaped structures; in the case of a
few of these structures, the shortest branch is simply removed. This process
might produce a disconnected backbone, at which point it is labelled
``severally branched''.

To gain experimental results, 45 digitised oil paintings from van Gogh and his
contemporaries were used, ranging in sizes and scales. A ground truth for the
brushstroke extraction was generated manually from 10 regions of these
paintings.

First, the system used to decide if a region is a brushstroke was evaluated on
these manually generated brushstrokes, obtaining a correct decision 95\% of the
time.

Then, the automatically extracted brushstrokes were compared to the manually
selected ones in terms of: validity (the percentage of extracted brushstrokes
which are in the ground truth) and detection (the percentage of the ground
truth strokes which were extracted).

Validity reached an average of 60\%, but ranged between 42\% and 83\% for
different paintings. Detection also reached and average of 60\%, but the lowest
was 22\% and only reached a high of 75\%.

Of course, the extracted brushstrokes also need analysis performed on them.
This is where the backbone calculation can also be used to extract features out
of the brushstroke, as well as considering the features of the brushstroke
itself.

Using this statistically, this can be used to authenticate paintings and also
begin to place them in different periods. However, the results from these are
in terms of $p$-values for each feature and do not consider multiple features.
However, it was found that the features which distinguish van Gogh from his
contemporaries are not the same as the ones which differentiate two different
periods in van Gogh's career.

These results are very promising, especially given that the brushstrokes were
compared to a ground truth. Of course it could be argued that the ten regions
selected were done so to show the best of the technique, but given the
visualisation of the brushstrokes, this seems unlikely.

It would have been interesting research to see how the threshold values for the
detection of a brushstroke; there is no investigation or discussion as to why
the high and low values were set where they were.

The same goes for $k$-means clustering; is there a better choice which provides
more accurate clusters, and therefore strokes, or was it chosen because of its
simplicity and the benefits gained from using a more complex one wouldn't be
worth the computational complexity?

One would have thought that a density-based clustering algorithm such as
DBSCAN\cite{kriegel11density} would have been more applicable to this problem
as it would have been able to cope with noise and may have provided a better
segment of the brushstrokes.

There is also the question of applicability. That is, this has been 
successfully applied to van Gogh and some of his contemporaries with oil
paintings, but would the technique still yield such good results is applied to
an artist who used a palette knife, etc.?

In conclusion, this paper presents a very useful method for extracting
brushstrokes and compares the process of doing so to a ground truth. There are
still some elements which could use further investigation, but on the whole
this is a very rounded piece of research.
